---
title: "sata_sort_Jon1"
author: "Jonathan Fielder"
date: '2022-11-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* Hypothesis - I expect to see a positive correlation between crime rate in a county and the average percentage of economically disadvantaged students. We will be looking to see if there is evidence in the data to support this hypothesis as well as looking for other possible correlations in the data such as: average grades in a county compared to economic disadvantage, economic disadvantage and types of crime (e.g. robbery vs. violent crime).

```{r imports}
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(moderndive)
library(readr)
Juris15 <- read_excel("Juris15.xlsx")
Juris16 <- read_excel("Juris16.xlsx")
Juris17 <- read_excel("Juris17.xlsx")
Juris18 <- read_excel("Juris18.xlsx")
Juris19 <- read_excel("Juris19.xlsx")
SchoolGrades15 <- read_excel("SchoolGrades15.xls")
SchoolGrades16 <- read_excel("SchoolGrades16.xls")
SchoolGrades17 <- read_excel("SchoolGrades17.xls")
SchoolGrades18 <- read_excel("SchoolGrades18.xls")
SchoolGrades19 <- read_excel("SchoolGrades19.xls")
```



```{r crime short}
crime15 <- Juris15
crime16 <- Juris16
crime17 <- Juris17
crime18 <- Juris18
crime19 <- Juris19

colnames(crime16) <- colnames(crime15)
colnames(crime17) <- colnames(crime15)
colnames(crime18) <- colnames(crime15)
colnames(crime19) <- colnames(crime15)


crime_15_19 <- rbind.data.frame(crime15, crime16, crime17, crime18, crime19)
```



```{r school sort}
grades15 <- SchoolGrades15[ -c(5:15,17,24)]
grades16 <- SchoolGrades16[ -c(5:19,21:23,30)]
grades17 <- SchoolGrades17[ -c(5:19,21:40,47)]
grades18 <- SchoolGrades18[ -c(5:19,21:41,48)]
grades19 <- SchoolGrades19[ -c(5:19,21:42)]


## oops more to get rid of
grades15 <- grades15[ -c(1,3,8,9)]
grades16 <- grades16[ -c(1,3,8,9)]
grades17 <- grades17[ -c(1,3,8,9)]
grades18 <- grades18[ -c(1,3,8,9)]
grades19 <- grades19[ -c(1,3,8,9)]



grades15 <- grades15 %>%
  mutate(Year = (2015))
grades16 <- grades16 %>%
  mutate(Year = (2016))
grades17 <- grades17 %>%
  mutate(Year = (2017))
grades18 <- grades18 %>%
  mutate(Year = (2018))
grades19 <- grades19 %>%
  mutate(Year = (2019))
  

colnames(grades16) <- colnames(grades15)
colnames(grades17) <- colnames(grades15)
colnames(grades18) <- colnames(grades15)
colnames(grades19) <- colnames(grades15)









## combine all years
grades15_19 <- rbind(grades15, grades16, grades17, grades18, grades19)

## rename to grade
colnames(grades15_19)[3] ="Grade"

## rename to county
colnames(grades15_19)[1] ="County"


## logic to numbers (100 yes 2 no) easier to switch to percent

grades15_19$`Charter School` <-ifelse(grades15_19$`Charter School` == "YES",100,0)
grades15_19$`Title 1` <-ifelse(grades15_19$`Title 1` == "YES",100,0)




## grades letters to numbers i.e. A=4 B=3 C=2 D=1 F=0
grades15_19 <- grades15_19 %>%
  mutate(Grade=ifelse(Grade == "A",4,Grade),
         Grade=ifelse(Grade == "B",3,Grade),
         Grade=ifelse(Grade == "C",2,Grade),
         Grade=ifelse(Grade == "D",1,Grade),
         Grade=ifelse(Grade == "F",0,Grade),
         Grade=ifelse(Grade == "I",5,Grade))


## removes incomplete grade schools for each year the data was not good for the school grade
grades15_19 <- subset(grades15_19, Grade != 5)

## remove incomplete minority percentages
colnames(grades15_19)[6] = "Minority"
grades15_19 <- grades15_19 %>%
mutate(Minority=ifelse(Minority == ".",200,Minority))
grades15_19 <- subset(grades15_19, Minority != 200)

## remove NA disadvantage
colnames(grades15_19)[7] = "Dis"
grades15_19 <- grades15_19 %>%
mutate(Dis=ifelse(Dis == ".",200,Dis))
grades15_19 <- subset(grades15_19, Dis != 200)


## summarize by county
countyGrades15_19 <- grades15_19 %>%
  group_by(County, Year) %>%
  summarize(`Title 1` = mean(as.numeric(`Title 1`)), 
            Grade = mean(as.numeric(Grade)),
            `Charter School` = mean(as.numeric(`Charter School`)),
            Minority = mean(as.numeric(Minority)),
            Disadvantaged = mean(as.numeric(Dis)))


## rename columns to reflect data
colnames(countyGrades15_19)[3] = "title_1_perc"
colnames(countyGrades15_19)[4] = "grade"
colnames(countyGrades15_19)[5] = "charter_perc"
colnames(countyGrades15_19)[6] = "minority_perc"
colnames(countyGrades15_19)[7] = "econ_disadv_perc"
  
```


countyGrades15_19 is the Melange


merger
```{r meging data frames}
county_data <- merge(countyGrades15_19,countyCrimes15_19,by=c("County","Year")) 
```


economic disadvantage vs crime rate per 100,000 per county per year
```{r econ vs crime per 100,000}
disadv_crime <- lm(econ_disadv_perc ~ county_data[[10]], data = county_data)
get_regression_table(disadv_crime)

ggplot(data = county_data, aes(y = econ_disadv_perc, x = county_data[[10]])) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Crime Rate Per 100,000", 
       y = "economic disadvantage", 
       title = "economic disadvantage vs crime rate per 100,000 per county per year") +
  theme_bw()

```
No correlation?



economic disadvantage vs grades per county per year
```{r econ vs grades}
disadv_crime <- lm(econ_disadv_perc ~ grade, data = county_data)
get_regression_table(disadv_crime)

ggplot(data = county_data, aes(y = econ_disadv_perc, x = grade)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Average Grades (0 = F, 4 = A)", 
       y = "economic disadvantage", 
       title = "economic disadvantage vs grades per county per year") +
  theme_bw()

```



grades vs crime rate per 100,000 per county per year
```{r grades vs crime per 100,000}
disadv_crime <- lm(grade ~ county_data[[10]], data = county_data)
get_regression_table(disadv_crime)

ggplot(data = county_data, aes(y = grade, x = county_data[[10]])) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Crime Rate Per 100,000", 
       y = "Average Grades (0 = F, 4 = A)", 
       title = "economic disadvantage vs grades per county per year") +
  theme_bw()

```

Slight correlation between grades in a county and the crime rate












